---
html_document:
  theme: united
author: "Minzhe Guo"
date: 'Compiled: `r Sys.Date()`'
output:
  html_document:
    df_print: paged
title: "Demonstration of cell type annotation using CellRef"
df_print: kable
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  tidy = TRUE,
  tidy.opts = list(width.cutoff = 95),
  message = FALSE,
  warning = FALSE,
  time_it = TRUE
)
```

```{r packages}
library(Seurat)
library(SeuratDisk)
library(ggplot2)
library(patchwork)
library(CellRef)
library(monocle3)
library(pheatmap)
```

First, let's create a Seurat object using the UMI count matrix (lung_endo.counts.rds) and cell information (lung_endo.cells.rds) of the demo data, which can be downloaded from the data folder of this CellRef github.

```{r load_data, warning=FALSE, message=FALSE}
cells = readRDS(file="./lung_endo.cells.rds")
counts = readRDS(file="./lung_endo.counts.rds")
query = CreateSeuratObject(counts=counts, meta.data = cells)
query = NormalizeData(query)

# batch correction of data from different donors
query = doDataIntegration(query, integration.batch ="DonorID",
                        method="Monocle3-mnn",
                        npcs=200, do.clustering = T)

# after integration, the method also perform clustering using leiden algorithm
g1 = DimPlot(query, reduction="umap",
             group.by="integrated_clusters", label=T)
g2 = DimPlot(query, reduction="umap",
             group.by="DonorID", label=F)
g2 = g2 + theme(legend.text = element_text(size=8))
g1+g2
```

Next, we use LungMAP Human Lung CellRef to annotate the cells in the Seurat object that we just created

```{r annotate, warning=FALSE, message=FALSE, fig.width=10, fig.height=4.5}
ref_ident = "celltype_level3"

ref = readRDS(file="./LungMAP_HumanLung_CellRef_Seed.v1.rds")

refmap.anchors <- FindTransferAnchors(
    reference = ref,
    query = query,
    normalization.method = "SCT",
    reference.reduction = "pca",
    dims = 1:dim(ref@reductions$pca@cell.embeddings)[2]
)

mapped <- MapQuery(anchorset = refmap.anchors, 
                   reference = ref, query = query,
                     refdata = list(cellref_celltype = ref_ident), 
					 reference.reduction = "pca", reduction.model = "umap")

```

We can prune the predictions with low confidence. By default, we use the mean-sd as the cutoff of prediction scores.

```{r }

score.mean = mean(mapped@meta.data$predicted.cellref_celltype.score)
score.sd = sd(mapped@meta.data$predicted.cellref_celltype.score)
score.cutoff = score.mean - score.sd

# We can see that while several non-Endo cell type were predicted, 
# they were all have very low confidence, 
# and will be pruned by our default cutoff (red) of the prediction score
g = ggplot(mapped@meta.data, aes(x=predicted.cellref_celltype, 
                                 y=predicted.cellref_celltype.score, 
                                 fill=predicted.cellref_celltype)) 
g = g + geom_boxplot(outlier.shape = NA)
g = g + geom_hline(yintercept = score.mean, color="black")
g = g + geom_hline(yintercept = score.cutoff, color="red")
g = g + theme_bw()
g = g + theme(axis.text.x = element_text(angle=90, vjust=0.5, hjust=1))
g

# pruned the predictions lower than the cutoff of prediction score
mapped@meta.data$predicted.cellref_celltype.pruned = as.character(mapped@meta.data$predicted.cellref_celltype)
mapped@meta.data$predicted.cellref_celltype.pruned[which(mapped@meta.data$predicted.cellref_celltype.score<score.cutoff)] = "pruned"
```

UMAP visualization of the cell clusters and CellRef annotations

```{r }

mapped.pruned = subset(mapped, predicted.cellref_celltype.pruned != "pruned")

g1 = DimPlot(mapped, reduction="umap",
             group.by="integrated_clusters", label=T)
g2 = DimPlot(mapped.pruned, reduction="umap",
             group.by="predicted.cellref_celltype.pruned", label=T)
g2 = g2 + theme(legend.text = element_text(size=8))
g1+g2

```

<details open>

<summary>

**Session Info**

</summary>

```{r}
sessionInfo()
```

</details>
